---
title: 'CS6190: Probabilistic Modeling Homework 3
MCMC'
author: "Gopal Menon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes:
- \usepackage{mathtools}
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{dirtytalk}
- \DeclareMathOperator{\Unif}{Unif}
- \DeclareMathOperator{\E}{E}
- \DeclareMathOperator{\Var}{Var}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\hrule

\begin{enumerate}

\item[1(a)] Gibbs Sampling of noisy image with Ising prior only.

The sampled images are shown for various values of $\alpha$ and $\beta$. The variance values shown should be disregarded as that part of the energy function is not used.

\end{enumerate}

```{r q1a, echo=FALSE, message=FALSE, fig.align='center'}
source("Gibbs Sampling Image Noise Reduction.r")
one_a_ising_prior_only()
```

\begin{enumerate}

\item[1(b)] Here are the posterior images for $10$ burn in and $20$ denoising iterations. The parameter values I used were $\alpha=0$, $\beta=1$ and $\sigma^2=1$.

\end{enumerate}

```{r q1b, echo=FALSE, message=FALSE, fig.align='center'}
full_posterior()
```

\begin{enumerate}

\item[1(c)] Final estimates for $\sigma^2=1$.

\end{enumerate}

```{r q1c, echo=FALSE, message=FALSE, fig.align='center'}
print_std_dev_estimates()
```

\begin{enumerate}

\item[2.] Say you are given data $(X,Y)$, with $X \in \mathbb{R}^d$ and $Y \in \{0,1\}$. The goal is to train a classifier that will predict an unknown class label $\tilde{y}$ from a new data point $\tilde{x}$. Consider the following model:

$$
\begin{aligned} 
Y &\sim Ber \left ( \frac{1}{1+e^{-X^T\beta}}\right ),\\
\beta &\sim N(0,\sigma^2I).
\end{aligned}\\
$$
This is a \textbf{Bayesian logistic regression} model. Your goal is to derive and implement a Hamiltonian Monte Carlo sampler for doing Bayesian inference on $\beta$.

\end{enumerate}

\begin{enumerate}

\item[(a)]Write down the formula for the unormalized posterior of $\beta|Y$, i.e.,

$$
\begin{aligned} 
p(\beta|y;x,\sigma) &\propto \prod_{i=1}^n p(y_i|\beta; x_i)p(\beta; \sigma)\\
&=\prod_{i=1}^n \left (\left [ \frac{1}{1+e^{-x_i^T\beta}} \right ]^{y_i} \left [1- \frac{1}{1+e^{-x_i^T\beta}} \right ]^{1-y_i}\right ) \frac{1}{\sqrt{(2\pi)^k |\sigma^2 I|} }e^{\left (-\frac{1}{2} \beta^T \left ( \sigma^2 I \right )^{-1} \beta\right)}
\end{aligned}\\
$$
\end{enumerate}

\begin{enumerate}

\item[(b)] Show that this posterior is proportional to $exp(-U(\beta))$, where
$$
\begin{aligned} 
U(\beta) = \sum_{i=1}^n (1-y_i) x_i^T \beta + \log(1+e^{-x_i^T\beta}) + \frac{1}{2\sigma^2}\|\beta\|^2
\end{aligned}\\
$$
$$
\begin{aligned} 
p(\beta|y;x,\sigma) &\propto \prod_{i=1}^n \left (\left [ \frac{1}{1+e^{-x_i^T\beta}} \right ]^{y_i} \left [1- \frac{1}{1+e^{-x_i^T\beta}} \right ]^{1-y_i} \right ) \frac{1}{\sqrt{(2\pi)^k |\sigma^2 I|} }e^{\left (-\frac{1}{2} \beta^T \left ( \sigma^2 I \right )^{-1} \beta\right)}\\
&= \prod_{i=1}^n \left ( \left [ \frac{1}{1+e^{-x_i^T\beta}} \right ]^{y_i} \left [\frac{e^{-x_i^T\beta}}{1+e^{-x_i^T\beta}} \right ]^{1-y_i} \right ) \frac{1}{\sqrt{(2\pi)^k |\sigma^2 I|} }e^{\left (-\frac{\|\beta\|^2}{2\sigma^2}   \right)}\\
\log \left ( p(\beta|y;x,\sigma) \right) &\propto -\sum_{i=1}^n y_i \log \left ( 1 + e^{-x_i^T\beta}\right ) + \sum_{i=1}^n (1-y_i).-x_i^T\beta - \sum_{i=1}^n (1 - y_i) \log \left ( 1 + e^{-x_i^T\beta}\right ) -\frac{1}{2\sigma^2}\|\beta\|^2\\
&= -\sum_{i=1}^n y_i \log \left ( 1 + e^{-x_i^T\beta}\right ) - \sum_{i=1}^n (1-y_i) x_i^T\beta - \sum_{i=1}^n \log \left ( 1 + e^{-x_i^T\beta}\right ) + \sum_{i=1}^n y_i \log \left ( 1 + e^{-x_i^T\beta}\right )  -\frac{\|\beta\|^2}{2\sigma^2}\\
&= - \sum_{i=1}^n (1-y_i) x_i^T\beta - \sum_{i=1}^n \log \left ( 1 + e^{-x_i^T\beta}\right )   -\frac{1}{2\sigma^2}\|\beta\|^2\\
\Rightarrow p(\beta|y;x,\sigma)  &\propto exp\left(-U(\beta)\right), \text{where}\\
U(\beta) &= \sum_{i=1}^n (1-y_i) x_i^T\beta + \sum_{i=1}^n \log \left ( 1 + e^{-x_i^T\beta}\right )   +\frac{1}{2\sigma^2}\|\beta\|^2
\end{aligned}\\
$$
\end{enumerate}

\begin{enumerate}

\item[2(e)] Trace plots of the $\beta$ sequence and histograms are shown below.

\end{enumerate}

```{r q2e, echo=FALSE, message=FALSE, fig.align='center'}
source("Iris Logistic Regression.r")
run_iris_logistic_regression()
```

\begin{enumerate}

\item[2(f)] A variance $\sigma^2$ value of $1$ was used, with a step size $\epsilon$ of $0.04$ and number of steps $L$ of $40$.

\end{enumerate}

```{r q2f, echo=FALSE, message=FALSE, fig.align='center'}
print_results()
```